---
title: "assigment_week3"
author: "Alinzon Simon"
date: "2024-09-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

##  1

Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

Response1: str_detect

```{r major_df, message=FALSE, warning=FALSE}

major_df <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv")

spec(major_df)

major_df |>
  filter( str_detect(Major,"DATA") | str_detect(Major,"STATISTICS") | str_detect(Major_Category,"DATA") | str_detect(Major_Category,"STATISTICS") )

```


##  2
Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"


Response1: Using Regex , str_squish and str_replace_all

```{r task2, message=FALSE, warning=FALSE}

task2 <-'[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"'

#Remove digits, [], additional line, and tab 
task2_result <- str_replace_all(task2, "\\d+|\\[|\\]|\\n|\\t", "")

#Remove extra whitespace from strings
task2_result <- str_squish(task2_result)

#Add a - between different elements in the vector
task2_result <- str_replace_all(task2_result, '" "', "-")

task2_result <- task2_result %>%
  str_replace_all('"', '') %>%    # Remove all quotes
  str_split("-", simplify = TRUE) %>%  # Split by dash
  as.vector()   # Convert to vector

#print the vector
print(task2_result)

```

##  3

Describe, in words, what these expressions will match:

## 3.1
(.)\1\1 , this will find the sequence of 3 characters. For example in the text, it replaced III for IV.

```{r 3.1, echo=FALSE}

str_replace_all("This is a version III", "(.)\\1\\1", "IV")


```

## 3.2
"(.)(.)\\2\\1" , this will find a sequence of two characters followed by its inverse, within double quotes. 
For example in the text, it replaced IXXI for "The Roman numeral IXXI".

```{r 3.2, echo=FALSE}

str_replace_all('"IXXI" is not valid', '"(.)(.)\\2\\1"', "The Roman numeral IXXI")


```

## 3.3
(..)\1 , this will find the sequence of 2 characters followed by same two characters. 
For example in the text, it replaced XIXI for "The Roman numeral IXXI".

```{r 3.3, echo=FALSE}

str_replace_all('XIXI is not valid', '(..)\\1', "The Roman numeral XIXI")


```


## 3.4
"(.).\\1.\\1" , this will find the sequence of 2 characters structure followed by the first character, within double quotes. 
For example in the text, it replaced XIXIX for "The Roman numeral IXXI".

```{r 3.4, echo=FALSE}

str_replace_all('"XIXIX" is not valid', '"(.).\\1.\\1"', "The Roman numeral XIXIX")



```


## 3.5
"(.)(.)(.).*\\3\\2\\1" , this will find the sequence of 3 characters structure followed by any content and then the same sequence reversed, within double quotes. 
For example in the text, it replaced "XIVXIV is not valid same as VIXVIX" for "The Roman numeral IXXI".

```{r 3.5, echo=FALSE}

str_replace_all('"XIVXIV is not valid same as VIXVIX"', '"(.)(.)(.).*\\3\\2\\1"', "String matches Regex")


```

# 4
Construct regular expressions to match words that:

## 4.1
Start and end with the same character : ^(.)[^\\1]*\\1$

```{r 4.1, echo=FALSE}

str_replace_all('. is not valid the The Roman numeral.', '^(.)[^\\1]*\\1$', "String matches Regex")


```


## 4.2
Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.): (.{2}).*\\1

```{r 4.2, echo=FALSE}

str_replace_all('XI XI ."', '(.{2}).*\\1', "String matches Regex")


```


## 4.3
Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.):(.).*\\1.*\\1

```{r 4.3, echo=FALSE}

str_replace_all('XIXIX ', '(.).*\\1.*\\1', "String matches Regex")


```