---
title: "Assigment 9 - Web APIs"
author: "Alinzon Simon"
date: "2024-10-25"
output: 
  html_document:
    code_folding: show
    theme:
      preset: "united"
      base_font:
        google: Merriweather
      heading_font:
        google: Proza Libre
    css: css/styles.css  
---


## Sentimental Analysis

In order to validate the result of the sentimental analysis I decided to get books that some key words are `sadness|melancholy|sorrow` in the subject and also another one with the word `happiness`.

The following packages were used:

- **gutenbergr:** package to export the books.

- **dplyr:** Enables flexible data frame manipulation.

- **stringr:**

- **DT:**

- **tidytext:**

- **tidyr:**

- **wordcloud:**

- **ggplot2:**

```{r main-libraries, message=FALSE, warning=FALSE}

install.packages("gutenbergr")
install.packages("stringr")
install.packages("wordcloud")

library(gutenbergr)
library(dplyr)
library(stringr)
library(DT)
library(tidytext)
library(tidyr)
library(ggplot2)
library(wordcloud)




```

### Identify two books

1) Searching for books with key words as `sadness|melancholy|sorrow`  

  - gutenberg_subjects : dataset that contains books
  
  - filter(): Select only the result that matches
  
  - str_detect(subject, regex("sadness|melancholy|sorrow", ignore_case = TRUE)) : detects any match with the regex rule; ignoring if it's lower or upper case

```{r search1}

sadness_subjects <- gutenberg_subjects |>
  filter(str_detect(subject, regex("sadness|melancholy|sorrow", ignore_case = TRUE)))

sadness_book_ids <- unique(sadness_subjects$gutenberg_id)

#additional validation to get only english
sadness_books_english <- gutenberg_metadata |>
  filter(gutenberg_id %in% sadness_book_ids, language == "en") |>
  select(gutenberg_id, title, author)


datatable(sadness_books_english,
          options = list(pageLength = 3,
          scrollX = TRUE))


```

2) Retrieve metadata for these books into a tibble frame

  - sadness_book_ids: int vector created in previous step with the book ids (gutenberg_id)
  
  - filter(gutenberg_id %in% sadness_book_ids): this part will select only the books that match with the gutenberg_id in the vector
  
  - select(gutenberg_id, title, author): columns that I will export for future use

```{r meta-data}

sadness_books_eng_data <- gutenberg_download(sadness_books_english$gutenberg_id)|>
  inner_join(sadness_books_english, by ="gutenberg_id")|>
  inner_join(sadness_subjects, by ="gutenberg_id")|>
  rename(book = title) 

# book 10800 was not found 

tidy_sadness_books_eng<- sadness_books_eng_data |>
  group_by(book)|>
  unnest_tokens(word, text) |>
  anti_join(stop_words)|>
  ungroup()|>
  mutate( linenumber = row_number())

datatable(tidy_sadness_books_eng,
          options = list(pageLength = 8,
          scrollX = TRUE))


```

Get information about number of words that matches the sadness sentiment

```{r sentiment_sadness}

nrc_sadness <- get_sentiments("nrc") |>
  #filter(sentiment == "sadness")
  filter(str_detect(sentiment, regex("sadness|melancholy|sorrow|negative", ignore_case = TRUE)))
  
tidy_sadness_books_eng_count <- tidy_sadness_books_eng |>
  group_by(book)|>
  inner_join(nrc_sadness) |>
  count(word, sort = TRUE)
  

datatable(tidy_sadness_books_eng_count,
          options = list(pageLength = 8,
          scrollX = TRUE))



```


I will use `get_sentiments("bing")`

```{r tidy_sadness_books_eng_data}


tidy_sadness_books_eng_sentiment <- tidy_sadness_books_eng %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)


datatable(tidy_sadness_books_eng_sentiment,
          options = list(pageLength = 8,
          scrollX = TRUE))


```


Now we will plot these sentiment scores across the plot trajectory for each book. For the aes we will use index and sentiment.

```{r tidy_sadness_books_eng_sentiment-plot}

# Wrap the book titles in the data frame
tidy_sadness_books_eng_sentiment <- tidy_sadness_books_eng_sentiment %>%
  mutate(book = str_wrap(book, width = 25))  # Adjust width as needed

# Plot with wrapped titles
ggplot(tidy_sadness_books_eng_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")+
  theme_minimal()


```

### Comparing the three sentiment dictionaries

- AFINN
- bing
- NRC


```{r comparing_3dictionaries}

afinn <- tidy_sadness_books_eng |> 
  inner_join(get_sentiments("afinn")) |> 
  group_by(index = linenumber %/% 80) |> 
  summarise(sentiment = sum(value)) |> 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  tidy_sadness_books_eng %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  tidy_sadness_books_eng %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)

datatable(afinn,
          options = list(pageLength = 8,
          scrollX = TRUE))

datatable(bing_and_nrc,
          options = list(pageLength = 8,
          scrollX = TRUE))



```

Now I will bend them together and visualize them 

```{r vizualization}

bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")+
  theme_minimal()


```

### Most common positive and negative words

By implementing count() here with arguments of both word and sentiment, we find out how much each word contributed to each sentiment.

```{r most_common_postive}

bing_word_counts <- tidy_sadness_books_eng |>
  inner_join(get_sentiments("bing")) |>
  count(word, sentiment, sort = TRUE) |>
  ungroup()

bing_word_counts

```

```{r most_common_postive-plot}

bing_word_counts |>
  group_by(sentiment) |>
  slice_max(n, n = 20) |> 
  ungroup() |>
  mutate(word = reorder(word, n)) |>
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)+
  theme_minimal()

```

### Wordclouds

We will use the tidy frame data into a wordclouds

- acast() should create a matrix where each row is a word, and each column is a sentiment (positive or negative). Check the matrix output to ensure it’s structured correctly.

```{r Wordclouds}

matrix_data <- bing_word_counts |>
  inner_join(get_sentiments("bing")) |>
  count(word, sentiment, sort = TRUE) |>
  acast(word ~ sentiment, value.var = "n", fill = 0)

head(matrix_data)

```

```{r Wordclouds}

bing_word_counts |>
  anti_join(stop_words) |>
  count(word) |>
  with(wordcloud(word, n, max.words = 40))


```

```{r  Wordclouds2}

library(reshape2)

bing_word_counts |>
  inner_join(get_sentiments("bing")) |>
  count(word, sentiment, sort = TRUE) |>
  acast(word ~ sentiment, value.var = "n", fill = 0) |>
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)


bing_word_counts |>
  inner_join(get_sentiments("bing")) |>
  count(word, sentiment, sort = TRUE) |>
  head(10)

```

