---
title: "Assignment Week5"
author: "Alinzon Simon"
date: "2024-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(knitr)
library(tidyr)
library(dplyr)
library(kableExtra)
library(DT) #The DT package interactive tables that support sorting, searching, and pagination
library(plotly) #interactive plots 
```

## 1)Loading data

In this assignment, the data was loaded from a CSV file. A column name was added for the first and second column in the data frame obtained from the CSV. 
Additionally, any missing values in the carrier column were populated with the previous record.

```{r load_data, echo=FALSE, warning=FALSE}
df_flights <- read_csv("https://raw.githubusercontent.com/alinsimon/data607/refs/heads/main/assigment5.csv",show_col_types = FALSE)
  

#I will rename the first columns
colnames(df_flights)[c(1,2)] <- c("carrier_name","flight_status")

#Before transformation the data source I will analyze the structure
#str(df_flights)

#Fill Down missing values for carrier_name
df_flights <- df_flights |>
  fill(carrier_name, .direction = "down") 

df_flights
```


## 2) Now we will identify the variables and observations

- The first column contains the destination names, and all destinations start with a capital letter. To extract this data, we will use a regular expression ^[A-Z]
- The corresponding numeric values associated with each destination will be loaded into a separate column named number_of_flights.
To do the transformation I will use pivot_longer, so I can change the data from wide to long.


```{r pivot_longer}

df_flights <- df_flights %>%
  pivot_longer(
    #Column names that start with a capital letter will become observations for a new column state
    cols = colnames(df_flights)[grep("^[A-Z]", colnames(df_flights))], 
    names_to = "destination",
    values_to = "number_of_flights",
    values_drop_na = TRUE
  )

#to use this function you need to install DT 

datatable(df_flights)

```

### 3) Perform analysis to compare the arrival delays for the two airlines.
A statistical summary was created to analyze the data. The carrier with the highest average number of delayed flights is AMMEST, while the carrier with the most on-time flights is AMMWEST. Additionally, AMWEST has a greater total number of flights compared to Alaska.

```{r statistical-summary}

summary_table <- df_flights |>
  group_by(carrier_name, flight_status) |>
  summarise(
    Mean = round(mean(number_of_flights), 1),
    Median = median(number_of_flights),
    SD = round(sd(number_of_flights), 1),
    Min = min(number_of_flights),
    Max = max(number_of_flights),
    Count = sum(number_of_flights),
    .groups = "drop"
  )

# Display the summary table using kable
summary_table  |>
  kable(caption = "Statistical Summary of Flights by Carrier and Flight Status")  |>
  kable_styling(full_width = F, position = "center")


```

```{r interactive-boxplot}


flights_box_plot <- ggplot(df_flights, aes(x = carrier_name, y = number_of_flights, fill = flight_status)) +
  geom_boxplot(outlier.shape = 21,  position = position_dodge(width = 0.8)) +
  facet_wrap(~ flight_status, nrow = 1) +  # Display flight_status in a single row of facets
   stat_summary(fun = mean, aes(label = round(..y.., 1)), geom = "text", vjust = 0.1,  color = "blue",, size = 3) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 2,  fill = "yellow", 
               position = position_dodge(width = 0.8)) +
  scale_fill_manual(values = c("on time" = "darkgreen", "delayed" = "red")) +  # The color I will use for the category
  labs(title = "Number of Flights by Carrier and Flight Status",
       subtitle = "Comparison between On Time and Delayed Flights",
       x = "Carrier Name",
       y = "Number of Flights") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "top", 
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))


# Convert ggplot to an interactive plotly plot
interactive_flights_box_plot <- ggplotly(flights_box_plot)

# Display the interactive scatter plot
interactive_flights_box_plot



```

```{r interactive-plot}

flights_scatter_plot <- ggplot(df_flights, aes(x = carrier_name, y = number_of_flights, color = flight_status)) +
  geom_point(size = 3, alpha = 0.7) +
  scale_color_manual(values = c("on time" = "darkgreen", "delayed" = "red")) +  # Use scale_color_manual
  labs(title = "Scatter Plot of Number of Flights by Carrier and Flight Status",
       x = "Carrier Name",
       y = "Number of Flights") +
  theme_minimal()


# Convert ggplot to an interactive plotly plot
interactive_flights_scatter_plot <- ggplotly(flights_scatter_plot)

# Display the interactive scatter plot
interactive_flights_scatter_plot

```

