---
title: "Assignment Week5"
author: "Alinzon Simon"
date: "2024-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
library(DT) #The DT package interactive tables that support sorting, searching, and pagination
library(plotly) #interactive plots 
```

## 1)Loading data

In this assignment, the data was loaded from a CSV file. A column name was added for the first and second column in the data frame obtained from the CSV. 
Additionally, any missing values in the carrier column were populated with the previous record.

```{r load_data, echo=TRUE, warning=FALSE}
df_flights <- read_csv("https://raw.githubusercontent.com/alinsimon/data607/refs/heads/main/assigment5.csv",show_col_types = FALSE)
  

#I will rename the first columns
colnames(df_flights)[c(1,2)] <- c("carrier_name","flight_status")

#Before transformation the data source I will analyze the structure
#str(df_flights)

#Fill Down missing values for carrier_name
df_flights <- df_flights |>
  fill(carrier_name, .direction = "down") 

df_flights
```


## 2) Now we will identify the variables and observations

- The first column contains the destination names, and all destinations start with a capital letter. To extract this data, we will use a regular expression ^[A-Z]
- The corresponding numeric values associated with each destination will be loaded into a separate column named number_of_flights.
To do the transformation I will use pivot_longer, so I can change the data from wide to long.


```{r pivot_longer, echo=TRUE}

df_flights <- df_flights %>%
  pivot_longer(
    #Column names that start with a capital letter will become observations for a new column state
    cols = colnames(df_flights)[grep("^[A-Z]", colnames(df_flights))], 
    names_to = "destination",
    values_to = "number_of_flights",
    values_drop_na = TRUE
  )

#to use this function you need to install DT 

datatable(df_flights)

```

### 3) Perform analysis to compare the arrival delays for the two airlines.
A statistical summary was created to analyze the data. The carrier with the highest average number of delayed flights is AMMEST and with the most average on-time flights is AMMWEST. Additionally, AMWEST has a greater total number of flights compared to Alaska.

```{r statistical-summary, echo=TRUE}

summary_table <- df_flights |>
  group_by(carrier_name, flight_status) |>
  summarise(
    Mean = round(mean(number_of_flights), 1),
    Median = median(number_of_flights),
    SD = round(sd(number_of_flights), 1),
    Min = min(number_of_flights),
    Max = max(number_of_flights),
    Count = sum(number_of_flights)
  )

# Display the summary table using kable
summary_table  |>
  kable(caption = "Statistical Summary of Flights by Airline and Flight Status")  |>
  kable_styling(full_width = F, position = "center")


```
When comparing based on destination and status, we can determine that Alaska has a higher percentage of on-time flights in Los Angeles, Phoenix, San Diego, San Francisco, and Seattle. In contrast, AMWEST has the highest percentage of delayed flights.

```{r PercentangeofFlights-Destination, echo=TRUE}

summaryt1 <- df_flights |>
  group_by(carrier_name, destination, flight_status) |>
  summarize(total_flights = sum(number_of_flights)) |>
  mutate(per_occurence = round(total_flights / sum(total_flights) * 100, 2))

summaryt1  |>
  kable(caption = "Flights by Carrier and Destination")  |>
  kable_styling(full_width = F, position = "center")

```
```{r PercentangeofFlights-Destination2, echo=TRUE, message=FALSE, warning=FALSE}

perc_flight_des_plot <-ggplot(summaryt1, aes(x = destination, y = per_occurence, fill = carrier_name)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste(per_occurence,'%')), 
            position = position_dodge(width = 0.9), 
            vjust = 0.5, angle = 90, hjust =1, color="#7F7F7F") +  
  facet_wrap(~ flight_status) +
  labs(y = "Percentange of Flights", x = "Destination",title = "Flights by Airline and Destination",
       subtitle = "Comparison between status")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual('Airline', values=c('#003863','#007461'))


# Convert ggplot to an interactive plotly plot
interactive_flights_geom_bar <- ggplotly(perc_flight_des_plot)

# Display the interactive scatter plot
interactive_flights_geom_bar


```

Overall, when we calculate the proportion of flights that arrived on time, we can see that AMWEST had the highest percentage at 89.11% .

```{r df_flights-Summary, echo=TRUE}

summaryt2 <- df_flights |>
  group_by(carrier_name,flight_status) |>
  summarize(total_flights = sum(number_of_flights)) |>
  mutate(per_occurence = round(total_flights / sum(total_flights) * 100, 2))

summaryt2  |>
  kable(caption = "Flights by Carrier and Status")  |>
  kable_styling(full_width = F, position = "center")




```


```{r interactive-plot, echo=TRUE}

flights_geom_bar <- 
ggplot(summaryt2, aes(x = flight_status, y = per_occurence, fill = carrier_name)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste(per_occurence,'%')), 
            position = position_dodge(width = 1), 
            vjust = 0.5, angle = 90, hjust =1, color="#7F7F7F") +
  labs(y = "Percentange of Flights", x = "Airline", fill = "Carrier",title = "Flights by Status",
       subtitle = "Comparison between status")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual('Airline', values=c('#003863','#007461'))



# Convert ggplot to an interactive plotly plot
interactive_flights_scatter_plot <- ggplotly(flights_geom_bar)

# Display the interactive scatter plot
interactive_flights_scatter_plot

```

