---
title: "Project2"
author: "Alinzon Simon"
date: "2024-10-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
library(DT) #The DT package interactive tables that support sorting, searching, and pagination
library(plotly) #interactive plots 
library(scales)
```

## DataSet I : Employee Salaries 2023

Employee_Salaries_-_2023.csv, this dataset was provided by Crystal Quezada.
The data was filtered based on the 'Department of Police,' and I would like to compare the salaries based on gender. To better display the data, I will create a subcategory for salary ranges:

* Low: 0 - 30,000\n
* Medium: 30,001 - 60,000\n
* High: 60,001 - 100,000\n
* Very High: 100,001 and above.


```{r load_data, echo=TRUE}

employee_salaries <- read_csv("https://raw.githubusercontent.com/alinsimon/data607/refs/heads/main/Employee_Salaries_-_2023.csv",show_col_types = FALSE)

employee_salaries_dep_pol <- employee_salaries |>
  filter(Department_Name == "Department of Police") |>
  select(Division, Gender, Base_Salary) |>
  mutate(Gender = if_else(Gender == "F", "Female", 
                          if_else(Gender == "M", "Male", "Unknown")),
         Salary_Range = cut(Base_Salary, 
                            # Define salary ranges
                            breaks = c(0, 30000, 60000, 100000, Inf),  
                            labels = c("Low", "Medium", "High", "Very High"),  
                            # Range labels
                            right = FALSE))  # Left-closed intervals

datatable(employee_salaries_dep_pol)

```
```{r summary-salaries, echo=TRUE}


summary_table <- employee_salaries_dep_pol |>
  group_by(Gender, Salary_Range) |>
  summarise(
    mean_salary = round(mean(Base_Salary), 2),
    median_salary = median(Base_Salary),
    sd_salary = round(sd(Base_Salary), 2),
    min_salary = min(Base_Salary),
    max_salary = max(Base_Salary),
    n = n(),
    upper = mean_salary + sd_salary,
    lower = mean_salary - sd_salary
  )

# Display the summary table using kable
summary_table  |>
  kable(caption = "Statistical Summary of Salaries in the Department of Police ")  |>
  kable_styling(full_width = F, position = "center")


```


In conclusion, the bar plot shows that, on average, females earn higher salaries in the range of $0 to $30,000, while males earn higher salaries in the medium, high, and very high categories.

```{r summary-salaries-plot, echo=TRUE}

ggplot(summary_table, aes(x = Salary_Range, y = mean_salary, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = dollar(mean_salary)),  # Use dollar format for the labels
            position = position_dodge(width = 0.9),  # Adjust position to align with bars
            vjust = -0.5,  # Adjust vertical position of the text
            size = 3) +  # Text size
  labs(title = "Average Salary by Gender",
       x = "Salary Range",
       y = "Salary $") +
  theme_minimal()+
  scale_y_continuous(labels = dollar_format()) # we had to add library scales
```

The highest salary by gender for the low-income category belongs to females. For the medium-income category, it belongs to males, while for the high-income category, it again belongs to females. However, in the very high-income category, males have a significantly higher salary.

```{r summary-salaries-plot2, echo=TRUE}

ggplot(summary_table, aes(x = Salary_Range, y = max_salary, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = dollar(max_salary)),  # Use dollar format for the labels
            position = position_dodge(width = 0.9),  # Adjust position to align with bars
            vjust = -0.5,  # Adjust vertical position of the text
            size = 3) +  # Text size
  labs(title = "The Highest Salary by Gender",
       x = "Salary Range",
       y = "Salary $") +
  theme_minimal()+
  scale_y_continuous(labels = dollar_format()) # we had to add library scales
```

## DataSet II : Car Comparisons
"2024 FE Guide for DOE-release dates before 9-17-2024-no-sales -9-17-2024public.xlsx", this dataset was provided by Benjamin Wolin.

```{r load_data2, echo=TRUE, message=FALSE, warning=FALSE}

cars_df <- read_csv("https://raw.githubusercontent.com/alinsimon/data607/refs/heads/main/2024%20FE%20Guide%20for%20DOE-release%20dates%20before%209-17-2024-no-sales%20-9-17-2024public.csv",show_col_types = FALSE)

#We will fix the column names to remove spaces and make them lower case
colnames(cars_df) <- tolower(gsub("[ ()#]", "_", colnames(cars_df)))
colnames(cars_df) <- sub("^__|_$|_-_", "", colnames(cars_df))


#Now I will create a new data frame with the rows that I need for the Search
cars_df_current_year <- cars_df|>
                        filter(model_year == as.numeric(format(Sys.Date(), "%Y"))) |>
                        select (model_year, mfr_name, carline, eng_displ, cyl, transmission, city_fe__guide_conventional_fuel, hwy_fe__guide_conventional_fuel)

datatable(cars_df_current_year)


  
```

We will now conduct an analysis based on the City fuel efficiency and the number of cylinders, using the new data frame cars_df_current_year. First, we will generate summary statistics to calculate the average fuel efficiency by the number of cylinders in the city.

```{r City_fuel_efficiency, echo=TRUE}

summary_table <- cars_df_current_year |>
  group_by(cyl) |>
  summarise(
    mean_city_fe= round(mean(city_fe__guide_conventional_fuel), 2),
    median_city_fe = median(city_fe__guide_conventional_fuel),
    sd_city_fe = round(sd(city_fe__guide_conventional_fuel), 2),
    min_city_fe= min(city_fe__guide_conventional_fuel),
    max_city_fe = max(city_fe__guide_conventional_fuel),
    total_count = n(),
    upper = mean_city_fe + sd_city_fe,
    lower = mean_city_fe- sd_city_fe
  )
summary_table  |>
  kable(caption = "Average City Fuel Efficiency (MPG)")  |>
  kable_styling(full_width = F, position = "center")



```

The bar plot shows that cars with fewer cylinders tend to have higher MPG (miles per gallon) in city driving.

```{r City_fuel_efficiency_plot, echo=TRUE}


ggplot(summary_table, aes(x = factor(cyl), y = mean_city_fe, fill = factor(cyl))) +
  geom_bar(stat = "identity", position = "dodge", color ="darkblue") +
  geom_text(aes(label = paste(mean_city_fe,'MPG')), 
            position = position_dodge(width = 1), 
            vjust = 0.5, angle = 90, hjust =1, color="white") +
  labs(y = "City Fuel Efficiency (MPG)", x = "Number of Cylinders", fill = "Cylinder",title = "Average City Fuel Efficiency (MPG)",
       subtitle = "Comparison between Number of Cylinders")+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12))+
  scale_fill_brewer(palette = "Blues", direction = -1)




```
 
 We will generate summary statistics to calculate the average fuel efficiency by the number of cylinders on the highway. On average, cars with 3 cylinders provide better fuel efficiency.

```{r hw_fuel_efficiency, echo=TRUE}

summary_table2 <- cars_df_current_year |>
  group_by(cyl) |>
  summarise(
    mean_hwy_fe= round(mean(hwy_fe__guide_conventional_fuel), 2),
    median_hwy_fe = median(hwy_fe__guide_conventional_fuel),
    sd_hwy_fe = round(sd(hwy_fe__guide_conventional_fuel), 2),
    min_hwy_fe= min(hwy_fe__guide_conventional_fuel),
    max_hwy_fe = max(hwy_fe__guide_conventional_fuel),
    total_count = n(),
    upper = mean_hwy_fe + sd_hwy_fe,
    lower = mean_hwy_fe - sd_hwy_fe
  )
summary_table2  |>
  kable(caption = "Average City Fuel Highway (MPG)")  |>
  kable_styling(full_width = F, position = "center")



```

The bar plot shows that cars with fewer cylinders tend to have higher Highway Fuel Efficiency (MPG). This trend is consistent with City Fuel Efficiency as well.

```{r hwy_fuel_efficiency_plot, echo=TRUE}


ggplot(summary_table2, aes(x = factor(cyl), y = mean_hwy_fe, fill = factor(cyl))) +
  geom_bar(stat = "identity", position = "dodge", color ="darkblue") +
  geom_text(aes(label = paste(mean_hwy_fe,'MPG')), 
            position = position_dodge(width = 1), 
            vjust = 0.5, angle = 90, hjust =1, color="white") +
  labs(y = "Highway Fuel Efficiency (MPG)", x = "Number of Cylinders", fill = "Cylinder",title = "Average Highway Fuel Efficiency (MPG)",
       subtitle = "Comparison between Number of Cylinders")+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12))+
  scale_fill_brewer(palette = "Blues", direction = -1)




```

A new category is created in order to display a graphic with both summaries.
```{r combined_summaries}
#Merging for summary data frame and using one column for the average
combined_summary <- bind_rows(
  summary_table |> 
    select(cyl, mean_city_fe) |> 
    rename(mean_fe = mean_city_fe) |>
    mutate(fuel_type = "City"),
  summary_table2 |>
    select(cyl, mean_hwy_fe) |>
    rename(mean_fe = mean_hwy_fe) |>
    mutate(fuel_type = "Highway")
)


ggplot(combined_summary, aes(x = factor(cyl), y = mean_fe, fill = fuel_type)) +
  geom_bar(stat = "identity", position = "dodge", color = "darkblue") +
  geom_text(aes(label = paste(mean_fe, 'MPG')), 
            position = position_dodge(width = 0.9), 
            vjust = -0.5, color = "black", size = 2.5) +
  labs(y = "Fuel Efficiency (MPG)", 
       x = "Number of Cylinders", 
       fill = "Fuel Type", 
       title = "City vs. Highway Fuel Efficiency (MPG)", 
       subtitle = "Comparison by Number of Cylinders") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12)) +
  scale_fill_brewer(palette = "Blues")

```

## DataSet III : I resubmitted Assigment week 5