---
title: "Project1"
author: "Alinzon Simon"
date: "2024-09-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyr)
library(stringr)
library(readr)
```

## Load RawData

```{r load_data, echo=TRUE}
raw_data <- read_csv("https://raw.githubusercontent.com/alinsimon/data607/refs/heads/main/project1data.txt",show_col_types = FALSE)

#rename the column
colnames(raw_data)[1] <- "text_value"

raw_data

```

## Analyze Source File
After analyzing the source file, I realized that each player's information is enclosed between dashes. For any row that starts with a number, I can extract the PairID, name, and Total. For rows that begin with two capital letters, I can obtain the remaining information. Consequently, two DataFrames were created with unique IDs so that they can be joined to generate the report file.

```{r get_id_state, echo=TRUE, warning=FALSE}
raw_data_All <- raw_data |>
  filter( row_number()!=1 , row_number()!=2)

Newrow <- 1
raw_data_ID_Info_PartI <- data.frame(PairID = character(),
                              PlayerName = character(),
                              total = character(), 
                              Round1 = character(), 
                              Round2 = character(), 
                              Round3 = character(), 
                              Round4 = character(), 
                              Round5 = character(), 
                              Round6 = character(), 
                              Round7 = character())


raw_data_ID_Info_PartII <- data.frame(PairID = character(),
                              State = character(),
                              uscf_id = character(), 
                              Rtg_Pre = character(), 
                              Rtg_Post = character(), 
                              Number = character())


for(i in 1:nrow(raw_data_All)){
  #if the row doesnt have -- 
  if (!grepl("--", raw_data_All[i, 1])) {
    #if row starts with id
      if (grepl("^[0-9]", raw_data_All[i, 1])) {
           temp  <-  raw_data_All[i, 1]|>
                                 separate(text_value, 
                                 into = c('PairID','PlayerName','Total','Round1',
                                 'Round2','Round3','Round4','Round5','Round6','Round7')
                                 , sep = "\\|", fill = "right")|>
                                 mutate(PairID = str_replace_all(PairID, " ", ""))
           
           raw_data_ID_Info_PartI<-bind_rows(raw_data_ID_Info_PartI,temp )
           PairID <- temp[1,'PairID'] # I will keep the ID so I can use it for the next row
           
      }else if (grepl("^[A-Z].{2}", raw_data_All[i, 1])){
          
           temp1  <-  raw_data_All[i, 1]|>
                                 separate(text_value, 
                                 into = c('State','uscf_id','Rtg_Pre','Rtg_Post',
                                 'Number')
                                 , sep = "[|]|[/]|[>]", fill = "right")|>
                                 mutate(across(everything(), str_trim)) |>
                                 mutate(Rtg_Pre = str_replace_all(Rtg_Pre, "[R: ]|[-]", ""))|>
                                 mutate(Number = str_replace_all(Number, "N:", ""))
           temp1[['PairID']] <- NA
           temp1[1,"PairID"] <- PairID
           
           raw_data_ID_Info_PartII<-bind_rows(raw_data_ID_Info_PartII,temp1 )
        }
  }
  
}
```

## Merge both DataFrames
In this part, I will combine the two previous DataFrames based on the PairID key. Additionally, I will convert the Round columns to numeric format to calculate the averages for the other players.

```{r mergedf, echo=TRUE}
#converting to Numeric
DF_Player <- inner_join(raw_data_ID_Info_PartI, raw_data_ID_Info_PartII, by = "PairID")|>
                                 mutate(across(everything(), str_trim))|>
                                  mutate(across(starts_with("Round"), 
                                  ~ str_replace_all(., " ", "")))|>
              mutate(across(starts_with("Round"), 
              ~ str_replace_all(., "[A-Z]", "")))|>
              mutate(across(starts_with("Round"), as.numeric))|>
              mutate(PairID = as.numeric(as.character(raw_data_ID_Info_PartI$PairID)))


DF_Player[['AveragePreChessRating']] <- NA

for(i in 1:nrow(DF_Player)){
  
  Previous_players <- DF_Player|>
  filter(  PairID == DF_Player[i, "Round1"] | PairID == DF_Player[i,"Round2"] |
           PairID == DF_Player[i,"Round3"] | PairID == DF_Player[i,"Round4"] |
           PairID == DF_Player[i,"Round5"] | PairID == DF_Player[i,"Round6"] |
           PairID == DF_Player[i,"Round7"] )|>
  select(Rtg_Pre)|>
  mutate(Rtg_Pre = as.numeric(str_remove(Rtg_Pre, "P.*")))
  
  if(nrow(Previous_players)>0){
    DF_Player[i,"AveragePreChessRating"] <-  mean(Previous_players[,"Rtg_Pre"],
                                                  na.rm = TRUE)|>
                                                  round()
  }
}

```

## Export result into a csv file

```{r closeconnection, echo=TRUE}
#Export into csv file

DF_Player_Export <- DF_Player|>
  select(PlayerName,State,Total,Rtg_Pre,AveragePreChessRating)

DF_Player_Export

write_csv(DF_Player_Export, "data/DF_Player.csv")

```